# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r echo = TRUE}
setwd("C:/Base/Education/MOOC/Johns Hopkins - Data Science Specialization/Reproducible Research/RepData_PeerAssessment1");
library("utils")
unzip("activity.zip", overwrite=TRUE)
data <- read.csv("activity.csv", header = TRUE, sep = ',')
```


## What is mean total number of steps taken per day?

```{r}
aggregate <- as.data.frame(tapply(data$steps, data$date, sum))
names(aggregate) <- c("total_steps")
```

```{r}
summary(aggregate)

plot (
    x    = as.Date(rownames(aggregate)),
    y    = aggregate$total_steps,
    type = "h",
    xlab = "Date",
    ylab = "Total Steps"
)
```


## What is the average daily activity pattern?
```{r}
avg <- as.data.frame(tapply(data$steps, data$date, mean))
names(avg) <- c("average_steps")
```

```{r}
which.max(avg[,1])

plot (
    x    = as.Date(rownames(avg)),
    y    = avg$average_steps,
    type = "l",
    xlab = "Interval Date",
    ylab = "Average Steps"
)
```


## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
summary(data$steps)[7]
```




## Are there differences in activity patterns between weekdays and weekends?
