# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r echo = TRUE}
setwd("C:/Base/Education/MOOC/Johns Hopkins - Data Science Specialization/Reproducible Research/RepData_PeerAssessment1");
library("utils")
if (!file.exists("activity.csv")) {
    unzip("activity.zip", overwrite=TRUE)
}
data_raw <- read.csv("activity.csv", header = TRUE, sep = ',')
data <- data_raw[complete.cases(data_raw),]
data$date <- as.Date(data$date)
head(data)
summary(data)
str(data)
```



## What is mean total number of steps taken per day?

```{r}
sum <- as.data.frame(tapply(data$steps, data$date, sum))
names(sum) <- c("total_steps")
```

```{r}
head(sum)
summary(sum)

plot (
    x    = as.Date(rownames(sum)),
    y    = sum$total_steps,
    type = "h",
    xlab = "Date",
    ylab = "Total Steps"
)

mean(sum$total_steps)
median(sum$total_steps)
```



## What is the average daily activity pattern?

```{r}
avg <- as.data.frame(tapply(data$steps, data$interval, mean))
names(avg) <- c("average_steps")
```

```{r}
head(avg)
summary(avg)

plot (
    x    = rownames(avg),
    y    = avg$average_steps,
    type = "l",
    xlab = "Interval",
    ylab = "Average Steps"
)

avg[which.max(avg[,1]),]
```



## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
summary(data_raw$steps)[7]
sum(!complete.cases(data_raw))
```

```{r}
data_fill <- data_raw
data_fill$date <- as.Date(data_fill$date)

# Disused line
#data_fill[which(is.na(data_fill$steps)),1] <- 1000

# Using this instead
for (i in 1:length(data_raw$steps)) {
  if (is.na(data_fill[i,1])) {
    data_fill[i,1] <- as.numeric(avg[(data_fill[i,3]==rownames(avg)),1])
  }  
}

head(data_fill)
```


```{r}
sum_fill <- as.data.frame(tapply(data_fill$steps, data_fill$date, sum))
names(sum_fill) <- c("total_steps")

plot (
    x    = as.Date(rownames(sum_fill)),
    y    = sum_fill$total_steps,
    type = "h",
    xlab = "Date",
    ylab = "Total Steps"
)

mean(sum_fill$total_steps)
median(sum_fill$total_steps)
```



## Are there differences in activity patterns between weekdays and weekends?

```{r}
data_diff <- data_fill
head(data_diff)

# The more intuitive way, but slower
for (i in 1:length(data_diff[, 2])) {
    if (weekdays(data_diff[i, 2])=="Saturday" | weekdays(data_diff[i, 2])=="Sunday") {
        data_diff$type[i]<-'weekend'
    } else {
        data_diff$type[i]<-'weekday'
    }
}

head(data_diff)

# A much faster way of doing it
weekday <- weekdays(data_diff$date, abbreviate = FALSE)
data_diff$type <- vector(length=length(data_diff$steps))
data_diff$type[weekday=="Sunday" | weekday=="Saturday"] <- "weekend"
data_diff$type[weekday!="Sunday" & weekday!="Saturday"] <- "weekday"
data_diff$type <- as.factor(data_diff$type)

head(data_diff)
```

```{r}
avg_diff <- aggregate(data_diff$steps, list(data_diff$interval, data_diff$type), mean)

head(avg_diff)

par(mfrow = c(2,1))

plot(x      = avg_diff[avg_diff$Group.2=="weekend",1], 
     y      = avg_diff[avg_diff$Group.2=="weekend",3], 
     xlab   = "Interval", 
     ylab   = "Average Steps", 
     type   ="l", 
     main   ="Weekend", 
     ylim   =c(0,250)
)

plot(x      = avg_diff[avg_diff$Group.2=="weekday",1], 
     y      = avg_diff[avg_diff$Group.2=="weekday",3], 
     xlab   = "Interval", 
     ylab   = "Average Steps", 
     type   ="l",
     main   ="Weekday",
     ylim   =c(0,250)
)
```
